---
title: "ggtukey"
author: "Ethan Bass"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggtukey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7.2,
  fig.height=4
)
```

```{r setup}
library(ggtukey)
library(ggplot2)
library(palmerpenguins)
data(penguins)
```

Simple boxplot with Tukey Letters:

```{r, fig.width=6}
boxplot_letters(data=penguins, x=species, y=body_mass_g, vjust=-0.5, hjust=-0.3, type="global")
```

# Faceting

Perhaps we want to break our penguin data by sex:

```{r}
boxplot_letters(data=penguins, x=species, y=body_mass_g, group=sex, vjust=-0.5)
```

Or look for sex differences within our three species:

```{r}
boxplot_letters(data=penguins, x=sex, y=body_mass_g, group=species, vjust=-0.5)
```

In these cases, there are two options for the statistics that may give different results. By default a `global` test is conducted where `TukeyHSD` is run on a full anova (\code{aov}) with class and year as interaction, returning all pairwise comparisons. If the comparisons are independent, `type="local"` can be selected to run independent Tukey tests within each group.

```{r}
boxplot_letters(data=penguins, x=species, y=body_mass_g, group=sex, type="local", vjust=-0.5)
```

Raw data can also be plotted in various ways using the `raw` argument to specify the desired geom for plotting raw data: e.g. `geom_point` (`points`), `geom_dotplot` (`dots`), `geom_jitter` (`jitter`).

```{r}
boxplot_letters(data=penguins, x = species, y = body_mass_g, group = sex, vjust = -0.5, raw = "jitter", alpha=0.6)
```
