---
title: "ggtukey"
author: "Ethan Bass"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggtukey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7.2,
  fig.height=4
)
```

```{r setup}
library(ggtukey)
library(ggplot2)
data(mpg)
```

Simple boxplot with Tukey Letters:

```{r, fig.width=6}
boxplot_letters(data=mpg, x=class, y=hwy,vjust=-0.5, hjust=-0.3, type="global")
```

Break it down by year:

```{r}
boxplot_letters(data=mpg, x=class, y=hwy, group=year, vjust=-0.5)
```

Here there are two options that may give slightly different results. By default a `global` test is conducted where `TukeyHSD` is run on a full anova with class and year as interaction, returning all pairwise comparisons. If the comparisons are independent, `type="local"` can be selected to run independent Tukey tests on each group.

```{r}
boxplot_letters(data=mpg, x = class, y = hwy, group = year, type="local", vjust = -0.5)
```

Raw data can also be plotted in various ways using the `raw` argument to specify the desired geom for plotting raw data: i.e. `geom_point` (`points`), `geom_dotplot` (`dots`), `geom_jitter` (`jitter`).

```{r}
boxplot_letters(data=mpg, x = class, y = hwy, group = year, type="local", vjust = -0.5, raw = "jitter", alpha=0.6)
```
